---
title: "Scrape_gutenberg"
author: "Ragnhild Sundsbak"
format: html
editor: visual
execute:
  eval: false
---

# Downloading a corpus from project gutenberg

This repository shows how to download a corpus of selected literature from gutenberg.org (Gitenberg 2026) using the Cran package "gutenbergr" (Bradford 2026).

This project creates a curpus of books with the subject "Political science"

When I write R-code, I always start with punching the absolute path, where I want my documents to be located. Then I ask R to confirm with getwd. The same result may be achieved through making a .Rproj, an R project and always klick on the .Rproj wile when you open the files. In that case, the getwd command will set your paths in the right place.

```{r}
#| label: setting-directory

# setwd("/Users/path/to/your/project")
# setwd may be used instead of an Rproj
getwd()
```

When we have installed all of the packages that needed, it is advisable to comment out the code with a \# like this:

```{r}
#| label: packages-libraries
# install.packages("gutenbergr")
#install.packages("devtools") 
#install.packages("usethis") 
# etc.

library(conflicted)
library(tidyverse)
library(gutenbergr) 
library(devtools) 
library(usethis)

conflict_scout()

conflicts_prefer(dplyr::filter()) 
conflicts_prefer(dplyr::lag())
```

The aim of the process is to prepare books as .md files for import into Obsidian vault.

```{r}
#| label: defining-the-corpus

political_science_subjects <- gutenberg_subjects |>
  filter(subject == "Political science")
```

Checking the results.

```{r}
View(political_science_subjects)
class(political_science_subjects)
```

Picking the relevant gutenberg_ids for the corpus, and checking out the vector.

```{r}
to_be_downloaded <- political_science_subjects$gutenberg_id

# Checking out the vector to_be_downloaded
View(to_be_downloaded)
class(to_be_downloaded)
length(to_be_downloaded)
```

The download might take some time, depending on how many books you have.

```{r}
# The download
Political_science_texts <- gutenberg_download(
  to_be_downloaded)

# If the job takes too long, and you get doubts, you may use 
# gutenberg_cache_clear_all()
```

## How many books/ gutenberg_ids do we have?

```{r}
Political_science_texts |> 
  summarise(
    number_of_rows = n(),
    number_of_books = n_distinct(gutenberg_id)
  )
```

## Lines pr book

```{r}
Political_science_texts |>
  count(gutenberg_id, name = "number_of_lines")
```

## Including the metadata

In gutenbergr, the metadata, and the supjects are split into different functions. Here, we start with the metadata.

```{r}
meta_p_sci <- gutenberg_metadata |>
  filter(gutenberg_id %in% to_be_downloaded)
```

## What does the metadata look like?

```{r}
glimpse(meta_p_sci) 
head(meta_p_sci)
```

## Subjects

We want all subjects for these books.

```{r}
subjects_p_sci_all <- gutenberg_subjects |>
  filter(gutenberg_id %in% to_be_downloaded)

glimpse(subjects_p_sci_all)
head(subjects_p_sci_all)
```

## Putting it together

This is supposed to solve the process. I had some help from ChatGPT (GPT UiO 2026). I plan to split this process up later, in order to facilitate understanding. But since it works, it will be kept here until I have a replacement. Readers may claim it is uneducational. In that case, I will appreciate their involvement.

```{r}
#| label: main-chunk
write_gutenberg_md <- function(book_id,
                               texts_df,
                               meta_df,
                               subjects_df,
                               out_dir = "books_md") {
  # Text lines for the book
  txt <- texts_df |>
    filter(gutenberg_id == book_id)
  
  if (nrow(txt) == 0) {
    warning("Ingen tekstlinjer for book_id = ", book_id)
    return(invisible(NULL))
  }
  
  # Metadata for the book
  meta <- meta_df |>
    filter(gutenberg_id == book_id)
  
  # All of the subjects for the book
  subs <- subjects_df |>
    filter(gutenberg_id == book_id)
  
  # Choose elements from metadata
  title    <- meta$title[1]
  author   <- meta$author[1]
  language <- meta$language[1] %||% "unknown"
  rights   <- meta$rights[1]    %||% ""
  
  subjects_str <- subs |>
    pull(subject) |>
    paste(collapse = "; ")
  
  # Merging the text lines
  full_text <- txt |>
    pull(text) |>
    paste(collapse = "\n")
  
  # Metadata-header
  header_lines <- c(
    paste0("# Title: ", title),
    paste0("# Author: ", author),
    paste0("# Gutenberg ID: ", book_id),
    paste0("# Language: ", language),
    paste0("# Rights: ", rights),
    paste0("# Subjects: ", subjects_str),
    "",
    "--- START OF TEXT ---",
    ""
  )
  
  md_lines <- c(header_lines, full_text)
  
  # Filename and folder
  if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)
  
  safe_title <- stringr::str_replace_all(title, "[^[:alnum:]]+", "_")
  file_name  <- file.path(out_dir, paste0(book_id, "_", safe_title, ".md"))
  
  writeLines(md_lines, con = file_name, useBytes = TRUE)
  message("Wrote ", file_name)
}
```

## Final step

This is the final step, and if everything works, the corpus will be downloaded to a folder called books_md

```{r}
purrr::walk(
  to_be_downloaded,
  ~ write_gutenberg_md(
    book_id     = .x,
    texts_df    = Political_science_texts,
    meta_df     = meta_p_sci,
    subjects_df = subjects_p_sci_all,  # alle subjects for disse bøkene
    out_dir     = "books_md"
  )
)
```

This was all, and you are now ready to make your own corpus.

## Literature:

Bradford, Jordan (2026). [Package "gutenbergr"](https://CRAN.R-project.org/package=gutenbergr).

GPT UiO (2026): [UiOs privacy friendly GPT chat](https://chatgpt.com/). The privacy friendly chat is for internal users. I have added the url for a similar service that is open to everyone.

Gutenberg, project (2026): [Project Gutenberg](https://gutenberg.org/).
